Experiences with Planning for Natural Language Generation

### 2008-07-20 trunk/Papers/experiences/conclusion.tex
revised for reviewer comments

In this paper, we have introduced two novel planning domains related
to problems in natural language generation: the sentence generation
problem and the GIVE navigation problem.  We have then reported on a
number of experiments in which we applied off-the-shelf planners to
these domains.  Our main results were that while modern planners do
pretty well in controlling the complexity of search, they suffer from
practical problems that limit their performance in unexpected places.
In particular, in both domains the grounding step that both planners
perform tends to take much more time than the search itself.

\todo{say something clever about the future}

>>>>

In this paper, we introduced two novel planning domains arising from
problems in natural language generation: the sentence generation
domain and the GIVE navigation domain. We also reported on the results
of a number of experiments in which we applied off-the-shelf planners
to these domains.

Our results were mixed. While modern planners do a pretty good job of
controlling the complexity of search, they also suffer from practical
problems that limit their performance in unexpected ways. In
particular, in both domains the grounding step performed by both
planners dominates the time it takes to perform the search itself. In
future work, it would be interesting to explore planners such as SHOP2
\cite{DBLP:journals/jair/NauAIKMWY03}, which do not perform the
initial grounding step.  We are also considering to propose some of
our domains as IPC benchmarks.

The NLG community's recent interest in planning presents a valuable
opportunity for planning researchers: provided some of the challenges
we have highlighted can be addressed, projects like GIVE (and other
NLG-inspired problem domains) offer a constructive platform for the
planning community to showcase their techniques to a wider
audience---and to improve the quality of their tools for real-world
planning tasks.  Furthermore, GIVE highlights planning-related
challenges such as plan execution monitoring and plan presentation
and summarisation and offers a platform on which such technologies can
be evaluated in experiments with human users.

\paragraph{Acknowledgments.} This work is a collaboration that arose
in the context of the ``planning and language'' group at the
University of Edinburgh.  We thank all members of this group,
especially Hector Geffner and Mark Steedman, for interesting
discussions.  This work was supported by the DFG Research Fellowship
``CRISP: Efficient integrated realization and microplanning'' and the
EU project XYZ PACO-PLUS.

### 2008-07-20 trunk/Papers/experiences/discussion.tex
revised for reviewer comments

The positive conclusion we can draw from the simple experiments reported
above is that modern planners are fairly good at controlling the search for
many of the moderately-sized NLG problems we looked at. This is
particularly true for FF in the sentence generation domain, which generates
nontrivial 14-word sentences in about two seconds. There is still room for
improvement, however. For instance, by comparison a special-purpose
algorithm for referring expressions could generate ten of these 14-word
sentences in a few milliseconds. Although search efficiency is not the main
performance metric in these domains, it nevertheless becomes a factor in
scenarios like Experiment~2, where the world is almost completely
unconstrained and (unsurprisingly) the search becomes much slower.

%The positive conclusion we can draw from the experiments reported
%above is that for the most part, modern planners are pretty good at
%controlling the search for the NLG-related inputs that we looked at.
%This is particularly true for FF in the sentence generation domain,
%which generates nontrivial 14-word sentences in about two seconds.
%There is still room for improvement (a special-purpose algorithm for
%referring expressions could generate ten of these 14 words in a few
%milliseconds), and the search becomes slower -- for obvious reasons --
%in the almost completely unconstrained world of Experiment 2, but
%search is not the main efficiency bottleneck.

The most restrictive bottleneck in the two domains we investigated is the
initial grounding step that many modern planners perform. While this may be
a good strategy for traditional planning benchmarks and IPC domains---where
plan length often dominates universe size, and an initial investment into
grounding pays off in saved instantiations during the search---this
approach is less effective in the domains we've considered. As our
experiments have shown, there are natural planning domains in which
relatively short plans must be computed in large universes. For instance,
in the case of the sentence generation domain, it is not unrealistic for
the universe to consist of thousands of domain individuals and tens of
thousands of actions, some of which require three domain individuals as
parameters. The common strategy of splitting the universe over several
types of individuals will not be very effective here.

%The most restrictive bottleneck in the two domains we have looked at
%in this paper is the initial grounding step that almost all modern
%planners perform.  Our conclusion from the experiments reported above
%is that this is a good strategy in domains where the typical plan
%length dominates the typical universe size, i.e.\ where the initial
%investment into grounding pays off in saved unifications during the
%search.  However, as we have shown, there are natural planning domains
%in which relatively short plans must be computed in large universes;
%it is not unrealistic for the universe in the sentence generation
%domain to consist of thousands of domain individuals and tens of
%thousands of actions, some of which take three domain individuals as
%parameters.  The common strategy of splitting the universe over
%several types of individuals will not be very effective here.

The recent trend in planning research has (rightfully) focused on the
development of algorithms that control search in sophisticated ways,
resulting in a host of planners that are more powerful and more successful
than their predecessors. Depending on the particular algorithm, however,
these planners often ground out sets of predicates and actions before
starting the search itself, which has a very pronounced effect on domains
such as those described above. Since our domains are not that unusual in
their structure and composition, we hope that the lessons learnt from our
experiences can help improve the performance of current systems. We believe
that such improvements are necessary if planning is to become a more mature
technology that can offer tools to a wider community of users. At the end
of the day, real-world users will care about the \emph{total} runtime of a
planner, and this is more than just the search time.
%We offer our larger problem instances as future challenges for the
%planning community.

%We believe that if planning is to become a mature technology which
%real-world users will believe in, it must find a way to deal with the
%grounding problem more gracefully than current systems do.  The
%research focus in developing planning algorithms has quite rightfully
%been on dealing with the search, and for this it is convenient --
%depending on the particular algorithm, necessary -- to ground out all
%predicates and actions before starting the search itself.  However, we
%think that the domains presented here are not that unusual in their
%ratio of plan length against universe size.  At the end of the day,
%real-world users will care about the \emph{total} runtime of a
%planner, and this is more than just the search.

@@

%By and large, it is a relatively pleasant experience to come as a
%``customer'' to the planning community: Thanks to the competitions, it
%is easy to identify and download a fast implementation, at least for
%Linux.  However, in the course of our experiments we found (and
%reported) bugs in both SGPLAN and FF.  And as our experiments have
%shown, there is no one silver bullet: Even planners as closely related
%as FF and SGPLAN can differ significantly in their performance on
%different domains.

>>>>

The positive conclusion we can draw from the simple experiments
reported above is that modern planners are fairly good at controlling
the search for many of the moderately-sized NLG problems we looked
at. This is particularly true for FF in the sentence generation
domain, which generates nontrivial 14-word sentences in about two
seconds. There is still room for improvement, however; special-purpose
algorithms generate much larger referring expressions in milliseconds
\cite{AreKolStr08}.  Although search efficiency is not the main
concern in these domains, it nevertheless becomes a factor in
scenarios like Experiment~2, where the world is almost completely
unconstrained and (unsurprisingly) the search becomes much slower.

The most restrictive bottleneck in the two domains we investigated is
the initial grounding step that many modern planners perform. While
this may be a good strategy for traditional planning benchmarks and
IPC domains---where plan length often dominates universe size, and an
initial investment into grounding pays off in saved instantiations
during the search---this approach is less effective in the domains
we've considered. As our experiments have shown, there are natural
planning domains in which relatively short plans must be computed in
large universes. For instance, in the case of the sentence generation
domain, it is not unrealistic to look for plans of length 20 over a
universe consisting of thousands of domain individuals and tens of
thousands of actions, some of which require three domain individuals
as parameters. The common strategy of splitting the universe over
several types of individuals will not be very effective here.

The recent trend in planning research has (rightfully) focused on the
development of algorithms that control search in sophisticated ways,
resulting in a host of planners that are more powerful and more
successful than their predecessors. However, the common strategy of
grounding out sets of predicates and actions in implementations of
these algorithms has a very pronounced effect on domains such as those
described above. Reachability analysis has the potential to help---we
could gain a lot by only computing reachable instances---, but this
too is typically done on the grounded problem. \todo{check if this is
  true} Since our domains are not that unusual in their structure and
composition, we hope that the lessons learnt from our experiences can
help improve the performance of current systems. We believe that such
improvements are necessary if planning is to become a more mature
technology that can offer tools to a wider community of users. At the
end of the day, real-world users will care about the \emph{total}
runtime of a planner, and this is more than just the search time.

@@

### 2008-07-20 trunk/Papers/experiences/domains.tex
revised for reviewer comments

In this section, we present our planning domains: the sentence
generation domain and the instruction giving domain.

@@

In practice, the process of determining referring expressions
typically interacts with the realization step and so it turns out to
be beneficial to perform both of these steps together. This was the
goal of the SPUD system \cite{Stone2003a}, which did this using a
top-down generation algorithm based on tree-adjoining grammars
\cite{joshi;etal1997} whose lexical entries were equipped with
semantic and pragmatic information. Unfortunately, SPUD suffered from
having to explore a huge search space, and had to resort to a
non-optimal greedy search strategy to retain reasonable efficiency.
To improve on the efficiency of SPUD,
\citeauthor{KolSto07}~\shortcite{KolSto07} translate the sentence
generation problem into a planning problem and use a planning
algorithm for generation.

@@

We compute this grammatical derivation top-down by starting with the
elementary tree for ``sleeps''. This tree satisfies the need to convey
the semantic information, but introduces a need to generate a noun
phrase (NP) for the subject; this NP must refer uniquely to the target
referent $r_1$. In a second step, we substitute the tree for ``the
rabbit'' into the open NP leaf, which makes the derivation
grammatically complete. Since there are two different individuals that
could be described as ``the rabbit''---technically, $r_2$ is still a
\emph{distractor} (i.e., based on the description ``the rabbit'', the
hearer might erroneously think that we're talking about $r_2$ and not
$r_1$)---we are still not finished. To complete the derivation, the
tree for ``white'' is added to the existing structure by the
\emph{adjunction} operation, making the derivation syntactically and
semantically complete.

@@

The process described above has clear parallels to planning: we manipulate
a state by applying actions in order to achieve a goal. We can make this
connection even more precise by translating the SPUD problem into a
planning problem. For instance, Figure~\ref{fig:white-rabbit-as-planning}
shows the corresponding PDDL actions for the above generation task, where
each action corresponds to an operation that adds a single elementary tree
to the derivation.  In each case, the first parameter of the action is a
node name in the derivation tree, and the remaining parameters stand for
the individuals to which the semantic roles will be instantiated.  The
syntactic preconditions and effects are encoded using $\mathsf{open}$ and
$\mathsf{canadjoin}$ literals; the status of each referring expression is
tracked using $\mathsf{distractor}$ literals. Notice that the action
effects contain terms of the form $\mathsf{subj}(u)$, which construct new
node names. In order to meet the syntactic requirements of PDDL, these
terms can be eliminated by estimating an upper bound $n$ for the plan
length, making $n$ copies of each action, ensuring that copy $i$ can only
be applied in step $i$, and replacing the term $\mathsf{subj}(u)$ in an
action copy by the constant $\mathsf{subj}_i$.

Once we've defined appropriate actions, we can solve the generation
problem as an ordinary planning problem. For instance, the following plan
solves our previous example:
%
\begin{enumerate}
\item $\mathsf{sleeps}(\mathsf{root}, r_1)$
\item $\mathsf{rabbit}(\mathsf{subj}(\mathsf{root}), r_1)$
\item $\mathsf{white}(\mathsf{subj}(\mathsf{root}), r_1)$
\end{enumerate}
%
Using this plan, the grammatical derivation in
Figure~\ref{fig:white-rabbit-sleeps-deriv}, and therefore the generated
sentence, can be systematically reconstructed. Thus, we can solve the
sentence generation problem via the detour through planning and bring
current search heuristics for planning to bear on generation.

@@

The object of the GIVE Challenge (``Generating Instructions in Virtual
Environments''; Koller et al.\ \citeyear{alexander07:_shared_task_propos})
is to build an NLG system which is able to produce natural-language
instructions which will guide a human user in performing some task in a
virtual environment. From an NLG perspective, GIVE makes for an
interesting challenge because it is a theory-neutral task that exercises
all components of an NLG system, and emphasizes the study of communication
in a (simulated) physical environment. It also has the advantage that the
user and the NLG system can be physically in different places, as long as
the 3D client and the NLG system are connected over a network. This makes
it possible to evaluate GIVE NLG systems on a large scale over the
Internet. The first GIVE evaluation will take place in late 2008;
currently eight research teams from five countries are working on
developing systems to participate in the challenge.\footnote{See the GIVE
 website for more details about this project:
 \url{homepages.inf.ed.ac.uk/v1akolle/proj/give/}.}

@@

The GIVE task comes with a number of interesting research challenges
both for the NLG and the planning communities.  On the NLG side, one
major problem is that the natural language instruction sequences can't
simply be obtained by verbalising each action instance in the plan in
turn.  For instance, the NLG system could generate an instruction
sequence starting with ``turn left; walk forward; press the button;
turn right; walk forward; walk forward; turn right; walk forward; walk
forward; turn left; walk forward''.  However, this would be clumsy and
boring, and it is to be expected that frustrated users will quickly
cancel an interaction with such a system, resulting in a bad
evaluation score.  Having just pressed the button, it would be much
better to say ``turn around and walk through the door''; i.e.,
multiple action instances should be merged into a single
instruction. Conversely, it may also be necessary to express a single
planning step with several instructions.  Having just entered the top
left room, it may be easier for the user to understand the instruction
``walk to the centre of the room; turn right; now press the green
button in front of you'' than the instruction ``press the green button
on the wall to your right''.  To plan the referring expression ``the
green button in front of you'' at a time when the button is not yet in
front of the user, the NLG system must keep track of the hypothetical
changes in what objects are visible to the user.  Thus the acts of
referring and instructing, and the modules for planning and NLG, must
be tightly integrated.

From a planning perspective, the use of planning in GIVE imposes very
strict runtime requirements on the planner: Planning must happen in
real time, because if the system takes too much time between deciding
to give an instruction and actually giving this instruction, the user
may have walked or turned away, thus making the instruction invalid.
Furthermore, plan execution monitoring also plays an important role in
the GIVE problem.  At a high level, the system needs to monitor a
user's actions and compare them against the generated instruction set
to determine if the user has correctly followed directions or not. In
the case of the latter, new instructions may have to be generated. In
practice, the situation can be quite complicated since the mental
state of the user is not known and so the system must observe the
user's actions in real time. For example, a user directed to ``turn
around and walk through the door'' may not necessarily perform these
actions to the letter, i.e., immediately turning 180 degrees and
proceeding directly to the door. Instead, the user might take a
roundabout route through the room, eventually exiting out the
door. Although the user's actions do not match the generated
instructions exactly, they meet the intended goal. The system must be
able to identify such ``equivalent'' plans and not immediately
generate new instructions as soon as the user's actions have gone off
course. Furthermore, a user can communicate certain intentions to the
system, both through action and inaction. For instance, the system
should infer that a user has failed to follow instructions if the user
exits a room when given a directive to ``walk to the centre of the
room''. However, the system should also make a similar conclusion if a
user simply does nothing when given the instruction.

>>>>

We begin by presenting our two planning domains: the sentence generation
domain and the instruction giving domain.

@@

In practice, the process of determining referring expressions typically
interacts with the realization step and so it turns out to be beneficial to
perform both of these steps together. This was the goal of the SPUD system
\cite{Stone2003a}, which performed this task using a top-down generation
algorithm based on tree-adjoining grammars \cite{joshi;etal1997} whose
lexical entries were equipped with semantic and pragmatic information.
Unfortunately, SPUD suffered from having to explore a huge search space,
and had to resort to a non-optimal greedy search strategy to retain
reasonable efficiency.  To improve on the efficiency of SPUD,
\citeauthor{KolSto07}~\shortcite{KolSto07} translate the sentence
generation problem into a planning problem and use a planning algorithm for
generation.

@@

We compute this grammatical derivation in a top-down manner, starting with
the elementary tree for ``sleeps''. This tree satisfies the need to convey
the semantic information, but introduces a need to generate a noun phrase
(NP) for the subject; this NP must refer uniquely to the target referent
$r_1$. In a second step, we substitute the tree for ``the rabbit'' into the
open NP leaf, which makes the derivation grammatically complete. Since
there are two different individuals that could be described as ``the
rabbit''---technically, $r_2$ is still a \emph{distractor} (i.e., based on
the description ``the rabbit'', the hearer might erroneously think that
we're talking about $r_2$ and not $r_1$)---we are still not finished. To
complete the derivation, the tree for ``white'' is added to the existing
structure by an \emph{adjunction} operation, making the derivation
syntactically and semantically complete.

@@

The process described above has clear parallels to planning: we
manipulate a state by applying actions in order to achieve a goal. We
can make this connection even more precise by translating the SPUD
problem into a planning problem.\footnote{See
  \url{http://code.google.com/p/crisp-nlg/} for the CRISP system, in
  which this conversion is implemented.}  For instance,
Figure~\ref{fig:white-rabbit-as-planning} shows the corresponding PDDL
actions for the above generation task, where each action corresponds
to an operation that adds a single elementary tree to the derivation.
In each case, the first parameter of the action is a node name in the
derivation tree, and the remaining parameters stand for the
individuals to which the semantic roles will be instantiated.  The
syntactic preconditions and effects are encoded using $\mathsf{open}$
and $\mathsf{canadjoin}$ literals; the status of each referring
expression is tracked using $\mathsf{distractor}$ literals. Notice
that the action effects contain terms of the form $\mathsf{subj}(u)$,
which construct new node names. In order to meet the syntactic
requirements of PDDL, these terms can be eliminated by estimating an
upper bound $n$ for the plan length, making $n$ copies of each action,
ensuring that copy $i$ can only be applied in step $i$, and replacing
the term $\mathsf{subj}(u)$ in an action copy by the constant
$\mathsf{subj}_i$.

Once the appropriate actions are defined, we can solve the generation
problem as an ordinary planning problem. For instance, the following plan
solves our previous example:
%
\begin{enumerate}
\item $\mathsf{sleeps}(\mathsf{root}, r_1)$
\item $\mathsf{rabbit}(\mathsf{subj}(\mathsf{root}), r_1)$
\item $\mathsf{white}(\mathsf{subj}(\mathsf{root}), r_1)$
\end{enumerate}
%
Using this plan, the grammatical derivation in
Figure~\ref{fig:white-rabbit-sleeps-deriv}, and therefore the
generated sentence ``the white rabbit sleeps'', can be systematically
reconstructed. Thus, we can solve the sentence generation problem via
the detour through planning and bring current search heuristics for
planning to bear on generation.

@@

The object of the GIVE Challenge (``Generating Instructions in Virtual
Environments''; Koller et al.\
\citeyear{alexander07:_shared_task_propos}) is to build an NLG system
which is able to produce natural-language instructions which will
guide a human user in performing some task in a virtual
environment. From an NLG perspective, GIVE makes for an interesting
challenge because it is a theory-neutral task that exercises all
components of an NLG system, and emphasizes the study of communication
in a (simulated) physical environment. It also has the advantage that
the user and the NLG system can be in physically different places, as
long as the 3D client and the NLG system are connected over a
network. This makes it possible to evaluate GIVE NLG systems on a
large scale over the Internet. The first GIVE evaluation will take
place in late 2008; currently eight research teams from five countries
are working on developing systems to participate in the
challenge.\footnote{See the GIVE website for more details about this
  project: \url{http://www.give-challenge.org/}.}

@@

The GIVE task comes with a number of interesting research challenges
both for the NLG and the planning communities.  On the NLG side, one
major problem is that the natural language instruction sequences can't
simply be obtained by verbalising each action instance in a plan in
turn.  For instance, the NLG system could generate an instruction
sequence starting with ``turn left; walk forward; press the button;
turn right; walk forward; walk forward; turn right; walk forward; walk
forward; turn left; walk forward''.  However, this would be clumsy and
boring, and it is to be expected that frustrated users will quickly
cancel an interaction with such a system, resulting in a bad
evaluation score.  Having just pressed the button, it would be much
better to say ``turn around and walk through the door''. That is,
multiple action instances should be merged into a single instruction;
plans must be \emph{summarised} in order to present them to the
user. Conversely, it may also be necessary to express a single
planning step with several instructions.  Having just entered the top
left room, it may be easier for the user to understand the instruction
``walk to the centre of the room; turn right; now press the green
button in front of you'' than the instruction ``press the green button
on the wall to your right''.  To plan the referring expression ``the
green button in front of you'' at a time when the button is not yet in
front of the user, the NLG system must keep track of the hypothetical
changes in what objects are visible to the user.  Thus the acts of
referring and instructing, and the modules for planning and NLG, must
be tightly integrated.

From a planning perspective, GIVE imposes very strict runtime requirements
on the planner and its associated modules: planning must happen in real
time and the system must respond to a user in a timely fashion.  If the
system takes too much time deliberating over an instruction to give, rather
than actually giving this instruction, the user may have walked or turned
away, thus making the instruction invalid.  Furthermore, plan execution
monitoring also plays an important role in the GIVE problem.  At a high
level, the system needs to monitor a user's actions and compare them
against the generated instruction set to determine if the user has
correctly followed directions or not. In the case of the latter, new
instructions may have to be generated. In practice, the situation can be
quite complicated since the mental state of the user is not known and so
the system must observe the user's actions in real time. For example, a
user directed to ``turn around and walk through the door'' may not
necessarily perform these actions to the letter, i.e., immediately turning
180 degrees and proceeding directly to the door. Instead, the user might
take a roundabout route through the room, eventually exiting out the door.
Although the user's actions do not match the generated instructions
exactly, they meet the intended goal. The system must be able to identify
such ``equivalent'' plans and not immediately generate new instructions as
soon as the user's actions have gone off course. Furthermore, a user can
communicate certain intentions to the system, both through action and
inaction. For instance, the system should infer that a user has failed to
follow instructions if the user exits a room when given a directive to
``walk to the centre of the room''. The system should also make a
similar conclusion if a user simply does nothing when given the
instruction.

%between deciding to give an instruction and

### 2008-07-20 trunk/Papers/experiences/experiments.tex
revised for reviewer comments

Results for the $h=20$ case, with $n$ ranging from $1$ to $40$, are shown
in Figure~\ref{fig:give-runtime-minimal}.  The most obvious result is that
FF is unable to solve any problems beyond $n=13$ on our experimentation
machine within the memory limit of 1 GB.  SGPLAN, on the other hand, solves
instances beyond $n=40$ without major problems.  The time spent on
grounding is not a major factor in either planner, probably because the
planners need more time to actually compute the plan---for instance, the
optimal plan for the problem $n=40$ has a length of about 1600 steps.

@@

Results for the $h=20$, $n=5$ case (with $w$ ranging from $1$ to $70$)
are shown in Figure~\ref{fig:give-runtime-junk}. As in Experiment~2,
FF again runs out of memory, this time at $w=17$. SGPLAN happily
solves inputs beyond $w=70$. However, unlike in Experiment 2, both
planners now spend a substantial proportion of their time on
grounding. In SGPLAN, this translates to a ``parsing time'' (which we
assume includes grounding) which grows from 180 ms to 21.7 seconds as
$w$ grows from $1$ to $75$. The rest of the runtime (which also
includes the search time) only grows from 400 ms to 2.3 seconds. This
difference is particularly dramatic given that the actual optimal plan
in each case is an identical plan of about 100 steps. The planning
times for these instances are also concerning since times over a
couple seconds negatively affect the overall response time of the
system, which must react in real time to user actions.

>>>>

Results for the $h=20$ case, with $n$ ranging from $1$ to $40$, are
shown in Figure~\ref{fig:give-runtime-minimal}.  The most obvious
result is that FF is unable to solve any problems beyond $n=13$ on our
experimentation machine within the memory limit of 1 GB.  SGPLAN, on
the other hand, solves instances beyond $n=40$ without major
problems.\footnote{Our ad-hoc implementation of GraphPlan is unable to
  find any of these plans.}  The time spent on grounding is not a
major factor in either planner, probably because the planners need
more time to actually compute the plan---for instance, the optimal
plan for the problem $n=40$ has a length of about 1600 steps.  As an
example, the plan for $n=h=1$ looks as follows:

\begin{enumerate}
\item $\mathsf{move}(\mathsf{pos\_1\_1},\mathsf{pos\_1\_2}, \mathsf{north})$
\item $\mathsf{manipulate}\textsf{-}\mathsf{u1}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{pos\_1\_2})$
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{north}, \mathsf{east})$
\item $\mathsf{move}(\mathsf{pos\_1\_2}, \mathsf{pos\_2\_2},
  \mathsf{east})$
\item $\mathsf{turn}\textsf{-}\mathsf{right}(\mathsf{east}, \mathsf{south})$
\item $\mathsf{move}(\mathsf{pos\_2\_2}, \mathsf{pos\_2\_1}, \mathsf{south})$
\item $\mathsf{manipulate}\textsf{-}\mathsf{l1}\textsf{-}\mathsf{off}\textsf{-}\mathsf{on}(\mathsf{pos\_2\_1})$
\end{enumerate}

@@

Results for the $h=20$, $n=5$ case (with $w$ ranging from $1$ to $70$)
are shown in Figure~\ref{fig:give-runtime-junk}. As in Experiment~2,
FF again runs out of memory, this time at $w=17$. SGPLAN happily
solves inputs beyond $w=70$. However, unlike in Experiment 2, both
planners now spend a substantial proportion of their time on
grounding. In SGPLAN, this translates to a ``parsing time'' (which we
assume includes grounding) which grows from 180 ms to 21.7 seconds as
$w$ grows from $1$ to $75$. The rest of the runtime (which also
includes the search time) only grows from 400 ms to 2.3 seconds. This
difference is particularly dramatic given that the actual optimal plan
in each case is an identical plan of about 100 steps (namely, the one
we would have found in Experiment 2 for $h=20$ and $n=5$). The
planning times for these instances are also concerning since times
over a couple seconds negatively affect the overall response time of
the system, which must react in real time to user actions.

### 2008-07-20 trunk/Papers/experiences/introduction.tex
revised for reviewer comments

Natural language generation (NLG; \citeauthor{reiter00building}
\citeyear{reiter00building}) is one of the major subfields of natural
language processing, concerned with computing natural language
sentences or texts that convey a given piece of information to the
user. Intuitively, this task can be viewed as a problem involving
actions, beliefs, and goals: an agent communicating with another agent
tries to change the mental state of the hearer by applying actions
which correspond to the utterance of words or sentences. This
characterization of NLG suggests obvious parallels to automated
planning---a view which has a long tradition in NLG
\cite{appelt:planning,young94dpocl}.  More recently, there has been a
resurgence of interest in applying modern planning techniques to NLG
\cite{Steedman-Petrick:07,KolSto07,benotti08b}.  While efficiency has
not always been the main focus of NLG, problems in NLG are often more
complex than their counterparts in other areas of natural language
processing (such as parsing \cite{KolStr02}), giving rise to
computationally challenging domains. These new approaches combine an
interest in planning as a modelling tool for natural language, with a
hope of improved efficiency by exploiting modern approaches and
algorithms.

@@

We will then discuss some of our experiences using off-the-shelf planners
in these domains. In particular, we explore the efficiency of FF
\cite{HoffmannNebel01} and SGPLAN
\cite{hsu06:_new_featur_in_sgplan_for}---planners that have been successful
on traditional benchmarks, and in the International Planning Competition
(IPC)---on a range of problem instances from our planning domains.  We also
compare these results against an ad-hoc Java implementation of GraphPlan
\cite{Blum1997}.  Our findings are mixed. On the one hand, it turns out
that modern planners handle some of the search problems that arise in NLG
quite easily (although large problem instances remain a challenge). On the
other hand, these same planners spend tremendous amounts of time on
preprocessing. For instance, FF spends 90\% of its runtime in the sentence
generation domain on grounding out literals and actions, and ends up slower
than our version of GraphPlan that avoids grounding. For domains like ours,
which are dominated by the number of actions and the universe size, rather
than the combinatorics of the search problem, this observation suggests
that the overall runtime of a planner could be improved by grounding more
selectively. We therefore offer these domains as challenges for the
planning community.

>>>>

Natural language generation (NLG; \citeauthor{reiter00building}
\citeyear{reiter00building}) is one of the major subfields of natural
language processing, concerned with computing natural language
sentences or texts that convey a given piece of information to the
user. Intuitively, this task can be viewed as a problem involving
actions, beliefs, and goals: an agent communicating with another agent
tries to change the mental state of the hearer by applying actions
which correspond to the utterance of words or sentences. This
characterisation of NLG suggests obvious parallels to automated
planning---a view which has a long tradition in NLG
\cite{appelt:planning,young94dpocl}.  More recently, there has been a
resurgence of interest in applying modern planning techniques to NLG
\cite{Steedman-Petrick:07,KolSto07,benotti08b}.  While efficiency has
not always been the main focus of NLG, problems in NLG are often more
complex than their counterparts in other areas of natural language
processing (such as parsing \cite{KolStr02}), giving rise to
computationally challenging domains. These new approaches combine an
interest in planning as a modelling tool for natural language, with a
hope of improved efficiency by exploiting modern approaches and
algorithms.

@@

We then discuss some of our experiences using off-the-shelf planners
in our two domains. In particular, we explore the efficiency of FF
\cite{HoffmannNebel01} and SGPLAN
\cite{hsu06:_new_featur_in_sgplan_for}---planners that are easily
available, support expressive preconditions and effects out of the
box, and have been successful on traditional benchmarks, and in the
International Planning Competition (IPC)---on a range of problem
instances from our planning domains.  We also compare these results
against an ad-hoc Java implementation of GraphPlan \cite{Blum1997}.
Our findings are mixed. On the one hand, it turns out that modern
planners handle some of the search problems that arise in NLG quite
easily (although large problem instances still remain a challenge).
On the other hand, these same planners spend tremendous amounts of
time on preprocessing. For instance, FF spends 90\% of its runtime in
the sentence generation domain on grounding out literals and actions,
and ends up slower than our version of GraphPlan that avoids
grounding. For domains like ours, which are dominated by the number of
actions and the universe size, rather than the combinatorics of the
search problem, this observation suggests that the overall runtime of
a planner could be improved by grounding more selectively. We
therefore offer these domains as challenges for the planning
community.
